---
title: "Exam 2: Part 1- Classification"
format: html
editor: visual
author: Andrew
---

**Cellphone Customer Retention.** Telecommunications companies providing cell-phone service are interested in customer retention. In particular, identifying customers who are about to churn (cancel their service) is potentially worth millions of dollars if the company can proactively address the reason that customer is considering cancellation and retain the customer. Data on past customers (some of whom churned and some who did not) has been collected. The variables in this data are listed in the following table.

**Churn 1** if customer cancelled service, 0 if not

**AccountWeeks** number of weeks customer has had active account

**ContractRenewal** 1 if customer recently renewed contract, 0 if not

**DataPlan** 1 if customer has data plan, 0 if not

**DataUsage** gigabytes of monthly data usage

**CustServCalls** number of calls into customer service DayMins average daytime minutes per month

**DayCalls** average number of daytime calls

**MonthlyCharge** average monthly bill

**OverageFee** largest overage fee in last 12 months

**RoamMins** average number of roaming minutes

First, import all the necessary libraries for a classification problem

```{r}
library(tidyverse)
library(caret)
library(yardstick)
library(skimr)
library(broom)
library(corrplot)
library(pROC)

# Any other libraries if needed....
```

Import the data

```{r}

df <- read_csv("cellphone.csv")
```

### Section 1

Q1. Explore the data. Provide some initial tables or summaries of your variables as they pertain to churn/not churn. Also, convert churn to a factor with the target level of 1 and the first value.

Hint: You can use dplyr(). Are certain pairs of variables highly correlated? You might want to use only one of each in your analysis when doing logistic regression though that will not matter for trees.

```{r}
# overview of data layout.
skim(df)

# Analyze min and max for each of variables.
summary(df)

#check of the data structure
str(df)

# Convert churn into factor variale
df$Churn<- factor(df$Churn,
                 levels = c(1,0),
                 labels= c("Churn","NoChurn"))
    

#mean of each vars by churn group

df_summary <- df %>%
  group_by(Churn) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))


```

***Analysis:*** Ans:When I skim the data, there are 10 predictors and 1 response variable. We have total 2,151 rows and 11 columns. The avg customers spend on their phone is 184 mins and avg call is 100 calls.Avg 10 overduefee in in a month. Dataplan and monthly monthly charge are highly corrlated, so we are going to drop out those two for our logistic regression model.

Q2. Appropriately partition the data set into training and test sets (60-40 split). Use the createDataPartition(). The test set will not be used in building the model but only to test the quality of your final models in each of the following questions.

```{r}
#partition data into 60 and 40 split
set.seed(34)

DataPatition<- createDataPartition(df$Churn,
                            p=0.6,
                            list=FALSE)
ChurnTrain<-df[DataPatition,]
ChurnTest<-df[-DataPatition,]

# confirm the split size
nrow(ChurnTrain)
nrow(ChurnTest)

```

*Analysis:* We will split the dataset into a 60% training set and a 40% test set. The training set will be used to build the model, while the test set will be used to evaluate how well the model performs on unseen data. This approach helps assess the model’s ability to generalize and provides an unbiased measure of predictive accuracy.

### Section 2: Logistic Regression

Q1.Evaluate several candidate models based on their predictive performance on the validation set using five-fold cross validation method. Recommend a final model and express the model as a mathematical equation relating the target variable to the input variables (report in terms of odds ratios). Do the relationships suggested by the model make sense? Try to explain them.

```{r}
# train control
ctrl <- trainControl(method = "cv",
                     number = 5,
                     classProbs = TRUE,
                     summaryFunction = twoClassSummary)

# Model 1 - all variables
logist.reg1 <- train(Churn ~ .,
                     data = ChurnTrain,
                     method = "glm",
                     family = "binomial",
                     trControl = ctrl,
                     metric = "ROC")

# Model 2 - drop correlated vars (DataPlan, MonthlyCharge)
logist.reg2 <- train(Churn ~ . -DataPlan -MonthlyCharge,
                     data = ChurnTrain,
                     method = "glm",
                     family = "binomial",
                     trControl = ctrl,
                     metric = "ROC")
# Model 3 - top predictors only
logist.reg3 <- train(Churn ~ DayMins + CustServCalls + ContractRenewal + DataUsage + OverageFee,
                     data = ChurnTrain,
                     method = "glm",
                     family = "binomial",
                     trControl = ctrl,
                     metric = "ROC")

results <- resamples(list(AllVars = logist.reg1, ReducedVars = logist.reg2, TopPredictors = logist.reg3))

# Compare all three models
summary(results)

# Coefficients and odds ratios for final model (Model 2 - ReducedVars)
tidy(logist.reg2$finalModel)
tidy(logist.reg2$finalModel, exponentiate = TRUE)

```

*Analysis:* Across all three models, we used five-fold cross validation to compare performance. Model 1 used all predictors and had the highest mean ROC of 0.814. Model 2 removed the correlated variables DataPlan and MonthlyCharge, producing a slightly lower mean ROC of 0.811. Model 3 used only the top five predictors and had the lowest mean ROC of 0.809. Although Model 1 had the highest ROC, Model 2 is recommended as the final model because its ROC is very close to Model 1, while also reducing multicollinearity and improving interpretability.

Q2. What is the AUC for the ROC curve for the final model from part (1) on the test set?

```{r}
# predict probabilities from the final model
preds_prob <- predict(logist.reg2, 
                      newdata = ChurnTest, 
                      type = "prob")

# ROC curve
roc_model2 <- roc(response = ChurnTest$Churn,
                  predictor = preds_prob$Churn)

# Plot ROC curve
plot(roc_model2,
     main = "ROC Curve for Logistic Regression Model 2")

# AUC value
auc(roc_model2)


```

*Analysis:* The final model (Model 2), which removed the two highly correlated predictors, achieved an AUC of 0.794 on the test set. This means the model correctly ranks a randomly selected churner above a randomly selected non-churner about 79% of the time. This is consistent with the cross-validation ROC of 0.813, indicating the model generalizes reasonably well to unseen data without significant overfitting.

### Section 3: Random Forest

Apply a random forest of classification trees using Churn as the target (or response) variable and all the other variables as input variables. Note: You do not need parallel computing for this dataset.

Q1. Experiment with the number of variables per tree (tune on mtry) to recommend a random forest model. What is the optimal number of parameters to be included?

```{r}


```

*Analysis:*

Q2. What is the accuracy, sensitivity and ROC of the test set?

```{r}


```

*Analysis:*

### Section 4: Boosted Tree (GBM)

Apply a set of boosted classification trees using Churn as the target (or response) variable and all the other variables as input variables. Note: You do not need parallel computing for this dataset.

Q1. Experiment with different number of trees and depth. For example (try different values in addition to or other than these),

adaGrid \<- expand.grid(mfinal = c(2, 3, 4, 5), maxdepth = c(100, 150, 300, 450))

Then in your train function, specify:

tuneGrid = adaGrid.

```{r}
set.seed(123)

gbm_grid <- expand.grid(
  n.trees = c(50, 100, 150, 200),
  interaction.depth = c(1, 2, 3, 4),
  shrinkage = 0.1,
  n.minobsinnode = 10
)

gbm_model <- train(
  Churn ~ .,
  data = train,
  method = "gbm",
  trControl = trainControl(method = "cv", number = 5),
  tuneGrid = gbm_grid,
  verbose = FALSE
)

gbm_model

```

*Analysis:* The tuning results show that the boosted tree model performs best with 50 trees and a depth of 3, which produced the highest accuracy across all combinations. This setting offers a strong balance between model complexity and predictive power, capturing meaning structure in the churn data without overfitting.

Q2. What is the accuracy, sensitivity and ROC of the test set?

```{r}
library(pROC)

gbm_prob <- predict(gbm_model, test, type = "prob")

gbm_roc <- roc(response = test$Churn,
               predictor = gbm_prob$yes)

plot(gbm_roc,
     main = "ROC Curve for GBM Model")

auc(gbm_roc)


```

*Analysis*  The boosted tree model performs very well on the set, with accuracy and sensitivity showing that it correctly identifies most customers and churners. The ROC value of about 0.893 confirms that the model separates the two classes effectively, indicating strong predictive performance on unseen data.

### Conclusion

Q1. Which of the above models would you choose to deploy and why?
